# Shortcut Functions

We have developed a LandTrendr JavaScript module that contains several shortcut functions to build LandTrendr input collections, run LandTrendr, and deal with the outputs. The module can be accessed from our public GEE repository. To use the module, you must first visit this URL: https://code.earthengine.google.com/?accept_repo=users/emaprlab/public. It will add the *users/emaprlab/public* repository to your GEE account. Once added, it can be found within the *Reader* permission group of your GEE scripts library. You'll find the LandTrendr module at *Modules/LandTrendr.js*.

To use the shortcut functions, you must import the *LandTrendr.js* module into your script using the following line - place it at the top of the script. 

```
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');
```

These functions use *USGS Landsat Surface Reflectance Tier 1* data sets. They mask clouds, cloud shadows, and snow from images using [CFMASK](https://www.sciencedirect.com/science/article/pii/S0034425714005069). Annual image composites are generated using a medoid approach: for a given image pixel, the medoid is the value for a given band that is numerically closest to the median of all corresponding pixels among images considered (all images between a provided annual data range). TM and ETM+ data are included an annual medoid compositing without alteration, but OLI image bands 2, 3, 4, 5, 6 and 7 are subset and transformed to the spectral properties of ETM+ bands 1, 2, 3, 4, 5 and 7, respectively, using slopes and intercepts from reduced major axis regressions reported in Roy et al  2016 Table 2.



## Functions

+ **[buildSRcollection]** - Builds an annual cloud and cloud shadow masked medoid composite of Landsat surface reflectance TM-equivalent bands 1,2,3,4,5,7. This collection can be useful outside of use by LandTrendr, but is also the base for creating the input collection for LandTrendr.
+ **[transformSRcollection]** - Transforms the images within an annual surface reflectance collection built by `buildSRcollection` to a list of provided indices or bands.
+ **[buildLTcollection]** - Builds a collection as input to LandTrendr. It will prepare a collection where the first band is the spectral index to base temporal segmentation on, and the subsequent bands will be fitted to segmentation structure of the segmentation index. 
+ **[runLT]** - Run LandTrendr given a set of parameters. This is a wrapper around `buildSRcollection` and `buildLTcollection` functions.
+ **[getSegmentData]** - Generates an array of information about spectral-temporal segments from the breakpoint vertices identified by LandTrendr.
+ **[getFittedData]** - Generates an annual band stack for a given index provided as *ftvList* indices to either *buildLTcollection* or *runLT*. It flattens the FTV array format to a band per year for a given FTV index.


### buildSRcollection

*Usage:*

`buildSRcollection(startYear, endYear, startDay, endDay, aoi)`

*Parameters:*

+ **startYear (Integer)**: The minimum year in the desired range of annual collection
+ **endYear (Integer)**: The maximum year in the desired range of annual collection
+ **startDay (String | month day formatted as 'mm-dd')**: The minimum date in the desired seasonal range over which to generate annual composite.
+ **endDay (String | month day formatted as 'mm-dd')**: The maximum date in the desired seasonal range over which to generate annual composite.
+ **aoi (Geometry)**: The area-of-interest over which to mosaic images

*Result:*

An ee.ImageCollection where each image represents the medoid of observations per TM-equivalent surface reflectance bands 1-5 and 7, for a given year (TODO: NEED TO LINK TO A SECTION DESCRIBING HOW WE BUILD ANNUAL SR COLLECTIONS). There will be as many images as there are years in the range inclusive of *startYear* and *endYear*. If a given year does not exist for the range, then a masked band will act as a filler. Similarly, if all observations of a given pixel within a year are masked because of cloud, cloud shadow, or snow, the pixel will be masked.

*Example:*

```
// load the LandTrendr.js module
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');

// define parameters
var startYear = 1985;
var endYear = 2017;
var startDay = '06-20';
var endDay = '09-20';
var aoi = ee.Geometry.Point(-122.8848, 43.7929);

// center and zoom the display in case outputs are to be mapped 
Map.centerObject(aoi,10);
Map.addLayer(aoi);

// apply LandTrendr.js functions
var annualSRcollection = ltgee.buildSRcollection(startYear, endYear, startDay, endDay, aoi);
print(annualSRcollection);
```



### transformSRcollection

*Usage:*

`transformSRcollection(srCollection, bandList)`

*Parameters:*

+ **srCollection (Image Collection)**: An annual surface reflectance collection generated by the `buildSRcollection` function.  
+ **bandList (List of strings)**: A list of one or more indices or bands from the list in the [Spectral index codes] section to be included in images composing the annual collection. Bands in the resulting images will be ordered and labeled as they are in this list.


*Result:*

An ee.ImageCollection that includes one image per year based on an image collection built by `buildSRcollection` function transformed to the indices provided in the `bandList` parameter.

*Example:*

```
// load the LandTrendr.js module
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');

// define parameters
var startYear = 1985;
var endYear = 2017;
var startDay = '06-20';
var endDay = '09-20';
var aoi = ee.Geometry.Point(-122.8848, 43.7929);
var bandList = ['NBR', 'NDVI', 'TCW'];

// center and zoom the display in case outputs are to be mapped 
Map.centerObject(aoi,10);
Map.addLayer(aoi);

// build annual surface reflectance collection (cloud and shadow masked medoid composite)
var annualSRcollection = ltgee.buildSRcollection(startYear, endYear, startDay, endDay, aoi);

// transform the annual surface reflectance bands to whatever is in the bandList variable
var indexCollection = ltgee.transformSRcollection(annualSRcollection, bandList);

// example of extracting year 2000 as an image from the collection
var year2000 = ee.Image(indexCollection.filterDate('2000-01-01','2000-12-31').first());
print(year2000);
```






### buildLTcollection

*Usage:*

`buildLTcollection(collection, index, ftvList)`

*Parameters:*

+ **collection (Image Collection)**: An annual surface reflectance collection generated by the `buildSRcollection` function. 
+ **index (String)**: The index from the list in the [Spectral index codes] section to be segmented by LandTrendr.
+ **ftvList (List of strings)**: A list of one or more indices from the list in the [Spectral index codes] section to be fitted to the segmentation of the `index` parameter. This parameter can also be set by and empty list `[]` which is valid and results in no FTV bands included in the LandTrendr output image.

*Result:*

An ee.ImageCollection where each image represents an assemblage of bands or indices to be segmented and fitted by LandTrendr. There will be as many images as there are years in the range inclusive of *startYear* and *endYear*. If a given year does not exist for the range, then a masked band will act as a filler. Similarly, if all observations of a given pixel within a year are masked because of cloud, cloud shadow, or snow, the pixel will be masked. The first band per image will be whatever spectral representation is defined by the `index` parameter - it will be oriented so that vegetation loss results in a positive spectral delta. Any following bands will be defined by the indices provided in the `ftvList` parameter, in the same order, and unmodified with regard to spectral delta orientation.

*Example:*

```
// load the LandTrendr.js module
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');

// define parameters
var startYear = 1985;
var endYear = 2017;
var startDay = '06-20';
var endDay = '09-20';
var aoi = ee.Geometry.Point(-122.8848, 43.7929);
var index = 'NBR';
var ftvList = ['NDVI', 'B4', 'B3'];

// center and zoom the display in case outputs are to be mapped 
Map.centerObject(aoi,10);
Map.addLayer(aoi);

// apply LandTrendr.js functions
var annualSRcollection = ltgee.buildSRcollection(startYear, endYear, startDay, endDay, aoi);
var annualLTcollection = ltgee.buildLTcollection(annualSRcollection, index, ftvList);
print(annualLTcollection)
```


### runLT

*Usage:*

`runLT(startYear, endYear, startDay, endDay, aoi, index, ftvList, runParams)`

*Parameters:*

+ **startYear (Integer)**: The minimum year in the desired range of annual collection
+ **endYear (Integer)**: The maximum year in the desired range of annual collection
+ **startDay (String | month day formatted as 'mm-dd')**: The minimum date in the desired seasonal range over which to generate annual composite.
+ **endDay (String | month day formatted as 'mm-dd')**: The maximum date in the desired seasonal range over which to generate annual composite.
+ **aoi (Geometry)**: The area-of-interest over which to mosaic images
+ **index (String)**: The index from the list in the [Spectral index codes] section to be segmented by LandTrendr.
+ **ftvList (List of strings)**: A list of one or more indices from the list in the [Spectral index codes] section to be fitted to the segmentation of the `index` parameter.
+ **runParams (Dictionary)**: A dictionary of parameters to control LandTrendr segmentation. You can find definitions for the dictionary keys and values in [LT parameters](./lt-gee-requirements.html#lt-parameters) section 


*Result:*

An array image described in the [LT-GEE Outputs](./lt-gee-outputs.html) section

*Example:*

```
// load the LandTrendr.js module
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');

// define parameters
var startYear = 1985;
var endYear = 2017;
var startDay = '06-20';
var endDay = '09-20';
var aoi = ee.Geometry.Point(-122.8848, 43.7929);
var index = 'NBR';
var ftvList = [];
var runParams = { 
  maxSegments:            6,
  spikeThreshold:         0.9,
  vertexCountOvershoot:   3,
  preventOneYearRecovery: true,
  recoveryThreshold:      0.25,
  pvalThreshold:          0.05,
  bestModelProportion:    0.75,
  minObservationsNeeded:  6
};

// center and zoom the display in case outputs are to be mapped 
Map.centerObject(aoi,10);
Map.addLayer(aoi);

// apply LandTrendr.js functions
var lt = ltgee.runLT(startYear, endYear, startDay, endDay, aoi, index, [], runParams);
print(lt)
```




### getSegmentData

*Usage:*

`getSegmentData(lt, index)`

*Parameters:*

+ **lt (Image Array)**: The image array that is the result of running LandTrendr (either `runLT()` or `ee.Algorithms.TemporalSegmentation.LandTrendr()`). 
+ **index (String)**: The index from the list in the [Spectral index codes] section to be segmented by LandTrendr.

*Result:*

An image array with dimensions: 7 (rows) x nSegments (cols). Each row describes an attribute of the segments idenitifed by LandTrendr per pixel time series. Each column represents a segment in the time series per pixel ordered from earliest to lastest in the series. 

+ Row 1: segment start year
+ Row 2: segment end year
+ Row 3: segment start value
+ Row 4: segment end value
+ Row 5: segment spectral delta
+ Row 6: segment duration
+ Row 7: segment rate of spectral change

To extract a row, use the `arraySlice` function. For example, if you wanted to work only with the segment start years, you would do this: `var segStartYr = arraySlice(segInfo, 0, 1)`. See the [Working with Outputs](./working-with-outputs.html#getting-segment-information) section for more information on handling these data.


*Example:*

```
// load the LandTrendr.js module
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');

// define parameters
var startYear = 1985;
var endYear = 2017;
var startDay = '06-20';
var endDay = '09-20';
var aoi = ee.Geometry.Point(-122.8848, 43.7929);
var index = 'NBR';
var ftvList = [];
var runParams = { 
  maxSegments:            6,
  spikeThreshold:         0.9,
  vertexCountOvershoot:   3,
  preventOneYearRecovery: true,
  recoveryThreshold:      0.25,
  pvalThreshold:          0.05,
  bestModelProportion:    0.75,
  minObservationsNeeded:  6
};

// center and zoom the display in case outputs are to be mapped
Map.centerObject(aoi,10);
Map.addLayer(aoi);

// apply LandTrendr.js functions
var lt = ltgee.runLT(startYear, endYear, startDay, endDay, aoi, index, ftvList, runParams);
var segInfo = ltgee.getSegmentData(lt, index)
print(segInfo)

```

### getFittedData

*Usage:*

`getFittedData(lt, startYear, endYear, index)`

*Parameters:*

+ **lt (Image Array)**: The image array that is the result of running LandTrendr (either `runLT()` or `ee.Algorithms.TemporalSegmentation.LandTrendr()`). 
+ **startYear (Integer)**: The minimum year in the desired range of annual collection
+ **endYear (Integer)**: The maximum year in the desired range of annual collection
+ **index (String)**: The index from the list in the [Spectral index codes] section to be segmented by LandTrendr.

*Result:*

An ee.Image representing fitted-to-vertex annual spectral data for whatever index was provided as the `index` parameter. There will be as many bands as there are years in the range inclusive of *startYear* and *endYear*.

*Example:*

```
// load the LandTrendr.js module
var ltgee = require('users/emaprlab/public:Modules/LandTrendr.js');

// define parameters
var startYear = 1985;
var endYear = 2017;
var startDay = '06-20';
var endDay = '09-20';
var aoi = ee.Geometry.Point(-122.8848, 43.7929);
var index = 'NBR';
var ftvList = ['NBR', 'NDVI'];
var runParams = { 
  maxSegments:            6,
  spikeThreshold:         0.9,
  vertexCountOvershoot:   3,
  preventOneYearRecovery: true,
  recoveryThreshold:      0.25,
  pvalThreshold:          0.05,
  bestModelProportion:    0.75,
  minObservationsNeeded:  6
};

// center and zoom the display in case outputs are to be mapped
Map.centerObject(aoi,10);
Map.addLayer(aoi);

// apply LandTrendr.js functions
var lt = ltgee.runLT(startYear, endYear, startDay, endDay, aoi, index, ftvList, runParams);
var nbrFTV = ltgee.getFittedData(lt, startYear, endYear, ftvList[0]);
var ndviFTV = ltgee.getFittedData(lt, startYear, endYear, ftvList[1]);
print(nbrFTV);
print(ndviFTV);
```










## Spectral index codes 

The LandTrendr.js functions library has a defined list of indices and bands that it will work with. Below in Table 9.1 is the list of 'index' codes to use and their common name. You should specify them as an all caps string.


```{r echo=F}
Code = c('NBR', 'NDVI', 'NDSI', 'NDMI', 'TCB', 'TCG', 'TCW', 'TCA', 'B1', 'B2', 'B3', 'B4', 'B5', 'B7')
Name = c('Normalized Burn Ratio', 'Normalized Difference Vegetation Index', 'Normalized Different Snow Index', 'Normalized Difference Moisture Index',
         'Tasseled-Cap Brightness', 'Tasseled-Cap Greenness', 'Tasseled-Cap Wetness', 'Tasseled-Cap Angle',
         'Thematic Mapper-equivalent Band 1','Thematic Mapper-equivalent Band 2','Thematic Mapper-equivalent Band 3','Thematic Mapper-equivalent Band 4',
         'Thematic Mapper-equivalent Band 5','Thematic Mapper-equivalent Band 7')

```




```{r echo=F, index-codes, tidy=FALSE}
knitr::kable(
  data.frame(Code, Name), 
  caption = 'Spectral Index Codes',
  booktabs = TRUE
)
```





